# Data Wrangle
Pulling data from files and reading them into R for analysis. 
```{r setup2, warning=F,message=F}
library(tidyverse)
library(tidyr)
library(ggthemes)
library(lubridate)

# Now that we have learned how to munge (manipulate) data
# and plot it, we will work on using these skills in new ways

#knitr::opts_knit$set(root.dir='..')
```


```{r dataread, warning=F,message=F}
####-----Reading in Data and Stacking it ----- ####
#Reading in files
files <- list.files('data',full.names=T)
##'data' will work if not inline printing


#Read in individual data files
ndmi <- read_csv(files[1]) %>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndmi')


ndsi <- read_csv(files[2]) %>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndsi')

ndvi <- read_csv(files[3])%>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndvi')

# Stack as a tidy dataset
full_long <- rbind(ndvi,ndmi,ndsi) %>%
  gather(key='site',value='value',-DateTime,-data) %>%
  filter(!is.na(value))
##new word is called pivot_longer

```




## Question (1) 

What is the correlation between NDVI and NDMI? - here I want you to
convert the full_long dataset in to a wide dataset using the 
function "spread" and then make a plot that shows the correlation s a
function of if the site was burned or not (x axis should be ndmi)
You should exclude winter months and focus on summer months
```{r}

full_wide<-spread(full_long, key='data', value='value')%>%
  mutate(month=month(DateTime))%>%
  filter( month %in% c(5,6,7,8,9))

ggplot(full_wide, aes(x=ndmi, y=ndvi,color=site ))+
  geom_point(shape=1) + 
  theme_few() + 
  scale_color_few() + 
  theme(legend.position=c(0.8,0.8))
print("There is a positive correlation between NDMI and NDVI for the summer months. This indicates that seasonal moisture impacts vegetation growth. ")

```
## Question (2) 

2) What is the correlation between average NDSI (normalized
 snow index) for January - April and average NDVI for June-August?
In other words, does the previous year's snow cover influence vegetation
 growth for the following summer?
```{r}

f_ndsi<-filter(full_long, data=='ndsi')%>%
  mutate(year=year(DateTime),month=month(DateTime))%>%
  filter(month%in% c(1,2,3,4))%>%
  group_by(year)%>%
  summarise(mean_ndsi=mean(value))

f_ndvi<-filter(full_long, data=='ndvi')%>%
  mutate(year=year(DateTime),month=month(DateTime))%>%
  filter(month%in% c(7,8,9))%>%
  group_by(year)%>%
  summarise(mean_ndvi=mean(value))

all_year<-inner_join(f_ndsi, f_ndvi, by='year')
  
ggplot(all_year, aes(x=mean_ndsi, y=mean_ndvi))+ 
  geom_point() + 
  theme_few() + 
  scale_color_few() + 
  theme(legend.position=c(0.8,0.8))
print("There is a weak positive correlation between NSDI and NDVI. This indicates that the previous years' snow cover influences the vegetation growth.")

```
## Question (3)

How is the snow effect from question 2 different between pre- and post-burn
and burned and unburned? 

```{r}
ndsi3<-filter(full_long, data=='ndsi')%>%
  mutate(year=year(DateTime),month=month(DateTime))%>%
  filter(month%in% c(1,2,3,4))%>%
  group_by(year,site)%>%
  summarise(mean_ndsi=mean(value))

ndvi3<-filter(full_long, data=='ndvi')%>%
  mutate(year=year(DateTime),month=month(DateTime))%>%
  filter(month%in% c(7,8,9))%>%
  group_by(year,site)%>%
  summarise(mean_ndvi=mean(value))

all_year3<-left_join(ndsi3, ndvi3)

ggplot(all_year3, aes(x=mean_ndsi, y=mean_ndvi, color=site))+ 
  geom_point() + 
  theme_few() + 
  scale_color_few()
message("The burned sites (blue) are more impacted by the previous years snow than the unburned (orange) sites")
```
## Question (4)

What month is the greenest month on average? 

```{r}
#plot it?
monthlygreen<-spread(full_long, key='data', value='value')%>%
  mutate(month=month(DateTime))%>%
  group_by( month)%>%
  summarise(meanNdvi=mean(ndvi, na.rm=TRUE))

ggplot(monthlygreen, aes(x=month, y=meanNdvi))+ 
  geom_point() + 
  theme_few() + 
  scale_color_few()+ 
  scale_x_continuous(name="Month Number", limits=c(1, 12), breaks=c(2,4,6,8,10,12))

Q4<-with(monthlygreen, month[which.max(meanNdvi)])
print(Q4)
message("On average the greenest month is the ", Q4, "th month.")
```


## Question (5) 

What month is the snowiest on average?
```{r}
snowy<-spread(full_long, key='data', value='value')%>%
  mutate(month=month(DateTime))%>%
  group_by( month)%>%
  summarise(meanNdsi=mean(ndsi, na.rm=TRUE))

ggplot(snowy, aes(x=month, y=meanNdsi))+ 
  geom_point() + 
  theme_few() + 
  scale_color_few()+ 
  scale_x_continuous(name="Month Number", limits=c(1, 12), breaks=c(2,4,6,8,10,12))

Q5<-with(snowy, month[which.max(meanNdsi)])
print(Q5)
message("On average the snowiest month is the ", Q5, "st month.")
```
