---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Functions and Iterations
Learning to quickly go through and pull data out of a dataset using functions and iterations to avoid repetitive code. 
```{r setup3, include=FALSE}
library(rvest)
library(tidyverse)
library(lubridate)
library(readxl)

```

```{r}
library(rvest)
site_url <- 'https://snowstudies.org/archived-data/'
#Read the web url
webpage <- read_html(site_url)

#Extract only weblinks and then the URLs!
links <- webpage %>%
  html_nodes('a') %>%
  .[grepl('forcing',.)] %>%
  html_attr('href')
message(links)
```
## Loop to read in data 
Download the meteorological data. Use the `download_file` and `str_split_fixed` commands to download the data and save it in your data folder. You can use a for loop or a map function. 
```{r}
splits <- str_split_fixed(links,'/',8)
#Keep only the 8th column
dataset <- splits[,8] 
#generate a file list for where the data goes
file_names <- paste0('data/',dataset)
for(i in 1:2){
  download.file(links[i],destfile=file_names[i])
}
downloaded <- file.exists(file_names)
list.dirs(path='3_snow_functions_iteration/data')


```
## Read in Function
Write a custom function to read in the data and append a site column to the data. 

```{r}

# this code grabs the variable names from the metadata pdf file
library(pdftools)
headers <- pdf_text('https://snowstudies.org/wp-content/uploads/2022/02/Serially-Complete-Metadata-text08.pdf') %>%
  readr::read_lines(.) %>%
  trimws(.) %>%
  str_split_fixed(.,'\\.',2) %>%
  .[,2] %>%
  .[1:26] %>%
  str_trim(side = "left")
headers #shows the headers of the dataset, useful for extracting columns. 
#function to read in data files

reader<- function(file){
  name = str_split_fixed(file,'/',2)[,2]%>%
  gsub('_Forcing_Data.txt','',.)
  df<-read_fwf(file)%>%
    select(c(1:11))
    names(df)<-headers[1:11]
    df<-df%>%
      mutate(site=name)
}
view(reader(file_names[1]))

```

## Map Function
Use the `map` function to read in both meteorological files. Display a summary of your tibble.
```{r}

met_files<- map_dfr(file_names, reader)
summary(met_files)

```
## Plotting data
Make a line plot of mean temp by year by site (using the `air temp [K]` variable). Is there anything suspicious in the plot? Adjust your filtering if needed.
```{r}
temperature<-met_files%>%
  group_by(year, site)%>%
  summarize(meanT = mean(`air temp [K]`))

ggplot(temperature, aes(x=year, y=meanT, color=site))+
  geom_line()+
  ggthemes::theme_few()+
  xlab("Year")+
  ylab("Mean Air Temperature (K)")

```
"Site SBB_SASP has a higher yearly mean temperature than site SBB_SBSP for all years in the dataset. There is a big jump in temperature from the first year to the following years. Further inspection (in Q6) is needed to determine the suspicious jump in data."

## Function to create multiple plots
Write a function that makes line plots of monthly average temperature at each site for a given year. Use a for loop to make these plots for 2005 to 2010. Are monthly average temperatures at the Senator Beck Study Plot ever warmer than the Snow Angel Study Plot?
Hint: https://ggplot2.tidyverse.org/reference/print.ggplot.html
```{r}

loopy<-function(data, xyear){
  #y<-as.vector(unique(data[c("year")]))
  met<-data%>%group_by(year, month , site)%>%
  summarize(meanT=mean(`air temp [K]`))%>%
    dplyr::filter(year==xyear)
  print(ggplot(met,aes(x=month, y=meanT, color=site))+
            geom_line()+ labs(y= "Mean Temperature (K)",x = "Year")+ facet_wrap(xyear))}
  
for (i in 2005:2010){
  loopy(met_files, i)
}
message("From 2005 to 2010 the monthly average temperatures at the Senator Beck Study Plot are never warmer than the Snow Angel Study Plot")

```
## Bonus
Make a plot of average daily precipitation by day of year (averaged across all available years). Color each site. 
```{r}
#coloring each site will not work with duplicate site data
met_files$second<-as.Date(with(met_files,paste(month,day,sep="-")),"%m-%d")%>%
format(., "%j")
  pre<-group_by(met_files,second)%>%
  summarize(meanPre = mean(`precip [kg m-2 s-1]`))
    names(pre)=c('Julian_Day', 'Mean_Precip')
ggplot(pre, aes(x=Julian_Day,y=Mean_Precip))+
  geom_point()+
  ggthemes::theme_few() +
  labs(x="Julian Day", y="Mean Precipitation (kg m-2 s-1)")


```
Bonus #2: Use a function and for loop to create yearly plots of precipitation by day of year. Color each site. 
```{r}

Ploopy<-function(data, xyear){
  #y<-as.vector(unique(data[c("year")]))
  met<-data%>%group_by(year, second , site)%>%
  summarize(meanP=mean(`precip [kg m-2 s-1]`))%>%
    dplyr::filter(year==xyear)
    colnames(met)[2]='Julian_Day'
  print(ggplot(met,aes(x=Julian_Day, y=meanP, color=site))+
            geom_point()+ labs(x="Julian Day", y="Mean Precipitation (kg m-2 s-1)") + facet_wrap(xyear))}
  
for (i in 2003:2010){
  Ploopy(met_files, i)
}

```
